---
title: "lab_08"
author: "James Hartner"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidycensus)
library('tidyverse')
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("9f3430d359e022a7620289f71ec793ba915a884f", overwrite = TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
ACS <- load_variables(2019, "acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
PPP_data <- read_rds('data/lab_08.rds')
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1**
Baltimore City is the jursidiction with the lowest median age, according to the ACS 2019 5-year estimates. 
```{r}
Question_one <- get_acs(geography= "county", variables="B01002_001", state= "MD", geometry = FALSE) 
```

**Q2** Which Maryland counties have a median age of at least 50?
**A2**
Talbot County and Worcester County have a median age of at least 50.
```{r}
Question_one %>%
  filter(estimate >= 50)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3**
Prince George's county had the highest percentage of 5-17 years olds that speak Spanish at home, with 25.60 percent. 
```{r}
spanish_speaking<- get_acs(geography= "county", variables="B16007_004", state= "MD", geometry = FALSE) 

spanish_speaking %>% select(GEOID, NAME, estimate) %>% rename(speaking_spanish=estimate) -> spanish_speaking
total_pop<-get_acs(geography= "county", variables="B16007_002", state= "MD", geometry = FALSE) 
total_pop %>% select(GEOID, NAME, estimate) %>% rename(population=estimate) -> total_pop

Joined_data <- total_pop %>% 
  left_join(spanish_speaking) %>%
  mutate(percentage=(speaking_spanish/population)*100) %>%
  arrange(desc(percentage)) 
print(Joined_data)
#B16007_004=5 to 17 speaking spanish
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**
The thing that stands out to me the most is how vastly different the top two counties are compared to the rest of the data. The top third had a very small population of 5,832 total 5-17 year olds. The top five counties are interesting because they are either in the highest populated counties in Maryland, or the lowest. Garrett County's number is very interesting because it is below a single percent at roughly .7 percent. The population  of 5-17 year olds of Garrett County (4078) is so much smaller than the top of Prince Georges County (143125). 
**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to creeate separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5**
The zip code was 60620 and the percentage of the population was 23.4196 %.
```{r}
NEW_PPP<-PPP_data %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  group_by(zip) %>%
  summarize(
    count=n()
  ) %>%
  arrange(desc(count))
entire_pop <-get_acs(geography= "county", variables="B01001_001", geometry = FALSE) 
entire_pop %>% select(GEOID, NAME, estimate) %>% rename(population=estimate) -> entire_pop
black_pop <-get_acs(geography= "county", variables="B02001_003", geometry = FALSE) 
black_pop %>% select(GEOID, NAME, estimate) %>% rename(black_pop=estimate) -> black_pop
joined_data_question_five <- black_pop %>%left_join(entire_pop) %>%
  mutate(percentage=(black_pop/population)*100) %>%
join_data_final <- NEW_PPP %>% inner_join(joined_data_question_five, by = c('zip'='GEOID')) %>%
  arrange(desc(count)) 
print(join_data_final)
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**
Nine out of the ten zip codes with the most removed applications are all located in Illinois, where Chicago is. When I was looking at a map with all of the zip code locations, nine out of the ten shared the common that they were in the Chicago. There was other zip code which was different. That zip code is located in Georgia. 
**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7**
I feel that cities and NAICS codes could be useful. I feel that these could show what industries are getting removed more compared to others. With cities, you could also focus on simply just one area and the population of that area to make some inferences in reporting of what happened to this removed loan. 