---
title: "Reverse Engineering Project"
author: "James Hartner, Daniel Lawall, Daniel Philipose"
date: "10/5/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction 

In this notebook, we are reverse engineering the story, ["Nearly 250 women have been fatally shot by police since 2015"](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Definitions
"Washington_Post_data.csv" is the data that we took from the github linked in the story. We decided to keep a consistent data set to look into (even though The Washington Post constantly updates this data), so our data was taken from Sept. 28. 

"WaPo_Data" is just "Washington_Post_data." Since we had multiple people working on this analysis, our data set is labled differently, but is the exact same data. 

'lubricate' is a package that we downloaded to help in question three. This helps modifies dates in our data to better analysis it. 

"drop_na()" is a function of R that drops the "NA" variables. For example, some of the victims had no registered race, so in order to clean the data, we had to drop those datapoints. The Washington Post did the same according to their sentence (sentence three). The sentence states, "of known races."

filter()- We'll use this if we want to see just one or more particular things from the data.

filter("" != "")- this takes everything else other than the piece we labeled. For example, say filter(race != "NA"), this takes out "NA" and keeps the rest of the races.

group_by - To group the variables we want.

summarize- This summarizes the data we pull.

n()- This counts how many datapoints are alike. For example, say we want to see how many Women were shot,we'll do group_by(gender) %>% summarize ( count=n()) to find how many women were fatally shot. 

mutate()- This adjusts the data set. 

mean()- This function takes the average of a piece of the data set. 

sum() takes the total amount of a piece of the data set. 

nrow() takes how many rows from the data set without needing to count. This is useful when we want to take the percentage of some piece of the data. 





## Packages
```{r}
install.packages('tidyverse')
install.packages('janitor')
install.packages('lubridate')
```
This story is from the Washington Post and was written by Marisa Iati, Jennifer Jenkins and Sommer Brugal
## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
Washington_Post_data<-read_csv("data/fatal-police-shootings-data.csv")
# Clean required data and prepare for analysis if needed. 
# The data was pretty cleaned. If it needed some additional cleaning, we put it in each respected question. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Paste in sentence to engineer here]
"About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed."

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]


Our findings were a little different because the story was posted over a year ago (Sept. 4, 2020) and the Washington Post have keep updating the data. In the data, which was taken Sept. 28, 2020, we found that since 2015, the percentage of women fatally shot by police who had mental health issues was higher at roughly 32 percent (95 of the 293 women). Men who had mental health issues was around 22 percent still ( 1398 of the 6303 men killed).



```{r}

# Put code to reverse engineer sentence here
#Data from Sept.28
women_fatally_shot<- Washington_Post_data %>%
filter(gender=="F")
women_fatally_shot %>%
group_by(signs_of_mental_illness) %>%
summarize(
total=n(),
percentage=total/293 #This took the percentage of victims who were women and had signs of mental illness
) %>%
arrange(desc(total))


#Men percentage
men_fatally_shot<-Washington_Post_data %>%
filter(gender=="M")
men_fatally_shot %>%
group_by(signs_of_mental_illness) %>%
summarize(
 count=n(),
 percentage=count/6298
) %>%
arrange(desc(count)) #same thing as the last set of code, except we wanted to see the percentage of men who had signs of mental illness and were killed. 

# What Prof. Mussenden did 
Mental_illness <- Washington_Post_data %>%
group_by(gender, signs_of_mental_illness) %>%
summarise(
   total=n()
) %>%
mutate(percentage=total/sum(total)) %>%
arrange(desc(total)) #This is an easier way to breakdown the dataset in terms of  percentage of people killed who had signs of mental illness. 


```



# Display results of code below this codeblock

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
"Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths."
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

We broke this statement into three parts in order to determine each of the listed  facts. The process was similar across all three parts, in which we either filtered or grouped by race and gender. In each part, we also used "mutate" to create a new column to determine the percentages. We were able to confirm all three of the facts in this statement: Black women account for less than 1% of fatal shootings where race is know, 20% of the women shot and 28% of the total number of unarmed deaths.

```{r}
overall_black_women <- Washington_Post_data %>%
  filter(race!="NA")
  group_by(race, gender) %>%
   summarise(
    count=n() #This is counting the genders that were registered from the victims. 
  ) %>%
  mutate(pct=(count/(nrow(Washington_Post_data)*100)))#This is taking the percentage of killings for each known race.
  

Black_women <- Washington_Post_data %>%
  filter(gender == "F") %>%
  filter(race != "NA" ) %>%
  group_by(race) %>%
  summarise(
    count=n()
  ) %>%
  mutate(pct_by_race=count/sum(count)) #Npw we are filtering the known races by gender to find how many Black women were fatally shot.

Unarmed_black_women <- Washington_Post_data %>%
  filter(gender == "F") %>%
  filter(armed == "unarmed") %>%
  filter(race != "NA")
  group_by(race) %>%
  summarise(
    count=n()
  ) %>%
  mutate(total_unarmed = sum(count)) %>%
  mutate(pct_by_race=count/total_unarmed) #Same thing as the previous code, except we're curious the percentage of black women that were unarmed.


```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
"The average age of women and men killed by police was 37, and about one-third were 25 to 34."
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

With this, we were able to find relatively the same results. For men, the  average age was 36.59 and for female, the average age was 37.87. However, if we are taking both men and women fatally shot combined, the average was roughly 36.66, which if we round that age to the nearest whole number, we'll get 37.

For the about "one-third" part of the sentence, our results show that roughly 31.35 percent of shooting victims are between the ages of 25 and 34, which is about one-third, or 33%. We believe that since the data is newer compared to the story, the percentage went down, and more data points of victims younger than 25 or older than 34 were added to the data set. 

```{r}
#Average for women
WaPo_data<- Washington_Post_data

age_fatally_shot <- WaPo_data %>%
  drop_na() %>%
  group_by(gender)%>%
  summarise(
    mean_age = mean(age)
  ) %>%
  print(age_fatally_shot)
Average_age <- WaPo_data %>%
  drop_na() %>%
  summarize(
    mean_age=mean(age)
  ) %>%
  print(Average_age)
 

#one-third 25 to 34
WaPo_data %>%
  filter(age >= "25" & age <= "34") %>%
  summarise(
    spec_count = n()
  ) %>%
mutate(pct=(spec_count/nrow(Washington_Post_data))*100)


# Display results of code below this codeblock

```

### Sentence 4
* **Sentence text**: [Paste in sentence to engineer here]
"The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings. Of those fatally shot every year, about 44 have been women."
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]
This roughly followed the same findings as Washington Post. We found that it was roughly 4% of the population killed were women from the police shootings. We also saw that if you take the average amount of women shot per year starting from 1-1-2015 (the first year the data started) to the last data point  (9/25/2021), it roughly equates to 44 women shot per year. If we took the average from the start of the first data date taken (1-2-2015) to the last, we would still see basically the same results. 
```{r}

#Ask Sean on if we should filter every year then take the average of the women dead or not for the 44 women every year
# Put code to reverse engineer sentence here
Question_3 <-Washington_Post_data %>%
filter(gender=="F")

#women_fatally_shot %>%
 # filter(gender=="F") %>%
  #summarise(
   #count = n()
  #) %>%
#mutate(pct=(count/5))
#Percentage of killings for women compared to men
women_fatally_shot %>%
  summarise(
   women_shot = n() #Here, we are counting how many women have been fatally shot by police. 
  ) %>%
  mutate(pct=(women_shot/nrow(Washington_Post_data))) #Here, we're are calculating the percentage of women shot from the whole data. 

years<-Washington_Post_data %>%
  mutate(
    CleanDate = as.POSIXct(date, format="%Y-%m-%d %H:%M:%S")
) #This converts the dates into a manageable format (founded in the class textbook). This will allow us to group the years and summarize the killings from each year as shown in the code below. 
  years %>%
  mutate(date = floor_date(date, "year")) %>% # We are creating a variable that groups the dates from their respective years, so, for example, everything from 2015 we'll all be summarized together. 
  filter(gender=="F") %>% #For this question, we want to see how many women were killed by police each year then take the average to see if it matches with what the Washington Post stated. 
  group_by(date) %>%
  summarise(
    total = n())%>%
  arrange(desc(total))
  years %>% #This will take the average 
  filter(gender=="F") %>%
  summarise(
    count=n(),
    average=count/7,#2015-2021 is considered seven years, since we are going from 01-01-2015 to 12-31-2021. 
   average_from_last_data_point =count/6.7315 #Note-2021 is not over, so technically, we should be dividing by the last datapoint registered as 9/25/2021/ so 6 years and 267 days which would be 6.7315 years. 
  )
```


### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

NOTE-We only had four sentences since one sentence was broken into two parts, but we kept it all in sentence 2 (see your comments from the first project_milestone)
```{r}
# Put code to reverse engineer sentence here
#We only had four sentences, since sentence 2 had two of the main points combined.
# Display results of code below this codeblock

```

-30-