---
title: "lab_06"
author: "James Hartner"
date: "10/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* [install Tabula](https://tabula.technology/)
* run existing code as directed (look for **Task**).
* write code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes in the lab_06 folder and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Get Our PDF

We'll be working with the [Maryland Statistical Handbook](https://planning.maryland.gov/MSDC/Documents/md-statistical-handbook.pdf) from the state Department of Planning. It's not tiny (44MB), and you'll want to download it to a place you'll remember (like your Downloads folder).

**Task** Download the file from the above link and save it to your Downloads folder.

## Setup Tabula

**Task** Start Tabula and load the PDF, then extract a table

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the Statistical Handbook file in your Downloads folder and click "open", and then click the "Import button" in Tabula. This will take a minute or two.

Let's go to page 30 of the PDF, Table 2A, "International Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". This is _net migration_, meaning it shows the total number of people moving to Maryland and its regions/jurisdictions _minus_ the total number of people leaving the same, divided into 12-month periods from July to July. In Tabula, draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV (it should be called `tabula-md-statistical-handbook.csv` by default) to your lab_06/data folder.

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)):

**Task** Write code to read the CSV file into a dataframe called `international_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores.

```{r}
# Write the code to load the CSV file here
international_migration <- read_csv("data/tabula-md-statistical-handbook-1.csv")
```

Add a column for the type of migration ("international") and populate it:

**Task** Write the code to add a `migration_type` column to the data and give it the value of "international"

```{r}
# Write the code to add the column here
international_migration<-read_csv("data/tabula-md-statistical-handbook-1.csv",skip=2,  col_names=c("county","July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total")) %>%
  filter(!is.na(county))
```

## Answer questions

**Q1** Which region and county/city _below_ the state level accounted for the largest percentage of international migration for the entire time period? You'll need to add and populate a column representing percent of total using `mutate`.
**A1** [write your answer here]
The region that accounted for the largest percentage was the Suburban Washington Region and the county was Montgomery. 
```{r}
# Write the code to produce your answer here
update_int_mig <- international_migration %>% 
  mutate(across(July_2011:total, parse_number)) %>%
  filter(county!= "State/Jurisdiction") %>%
  mutate(international_percentage= total/198996)
update_int_mig %>%
group_by(county,international_percentage) %>%
summarize(
  percentage=(international_percentage*100)
) %>%
arrange(desc(percentage))

  
```

**Q2** Write a sentence or two that describes the data you produced in A1. Try to convey the most important idea.
**A2** [write your answer here]
The data that was produced in A1 shows each region of Maryland and the percentage of international migrants, who moved to Maryland or left (negative number) in the past few years, in each region/county. This is important to see because it can help show the most popular regions in Maryland and could potentially lead to a story on why people are moving there or leaving there. It can also help paint a better picture of where a region can work on to have more people moving in instead of moving out.
**Q3** Which region & jurisdiction had the largest percentage change (either negative or positive) for international migration between July 2018 and July 2017? The formula for percentage change is easy to remember: (New-Old)/Old.
**A3** [write your answer here]
For Counties, Kent County had the highest percent change at 28.57 percent and the Lower Eastern Region had the best percent change for a region at -14.09 percent (rounded to -14.10). The largest percent change (negative though) for a region was the Southern Maryland Region at -58.46 percent.
```{r}
# Write the code to produce your answer here
percent_change<-update_int_mig %>%
  mutate(percentage_change=((July_2018-July_2017)/July_2017))
percent_change %>%
group_by(county, percentage_change) %>%
  summarize(
  percentage=percentage_change*100
) %>%
arrange(desc(percentage))
  


```

**Q4** What's your best guess as to why these changes occurred, and in those places in particular?
**A4** [write your answer here]
My best guest for Kent County on having the highest percent change of any county was due to small amount of people who moved internationally to there were (2017 had seven people moving there and 2018 had nine).

The Southern Maryland Region also saw a small number. In 2017, there were 260 people moving there and 2018 only saw 108. 

## Back to Tabula

**Task** Extract second table from Tabula

Let's go to page 31 of the PDF, Table 2B, "Domestic Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". In Tabula, hit the "Clear All Selections" button and then draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV to your lab_06/data folder. (and rename it to `tabula-md-statistical-handbook-domestic.csv`).

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner:

**Task** Write code to read the CSV file into a dataframe called `domestic_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores. Your work on the international file should serve as a guide.

```{r}
# Write the code to load the CSV file here
domestic_migration <- read_csv("data/tabula-md-statistical-handbook-domestic.csv")
```

Add a column for the type of migration ("domestic") and populate it:

**Task** Write the code to add a `migration_type` column to the data and populate it with the value of "domestic"

```{r}
# Write the code to add the column here
domestic_migration <- read_csv("data/tabula-md-statistical-handbook-domestic.csv",skip=2,  col_names=c("county","July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total")) %>%
  filter(!is.na(county))
```

## Answer questions
**Q5** Which Maryland individual jurisdiction saw the largest net decrease in domestic migration overall?
**A5** [write your answer here]
The Suburban Washington Region saw the largest net decrease in domestic migration overall. 
```{r}
# Write the code to produce your answer here
domestic_migration %>%
group_by(county, total) %>%
summarise(
  
) %>%
  arrange(total)


```

**Q6** How many regions & jurisdictions had net positive migration for July 2017, July 2018 and July 2019 (not just individually but all three together)?
**A6** [write your answer here]
There were 18 regions and jurisdictions that had net positive migration for all three years together. 
```{r}
# Write the code to produce your answer here
Net_positives <- domestic_migration %>%
  mutate(question_6_years= (July_2017+July_2018+July_2019)) 
Net_positives %>%
group_by(county,question_6_years) %>%
summarize(
  
) %>%
  arrange(desc(question_6_years)) %>%
filter(question_6_years > '0')
```

**Q7** How would you describe this data? Is there a county or region that stands out, and why?
**A7** [write your answer here]
I feel that Frederick is very interesting because it is the largest net positive over the years in domestic migration. Frederick County is located in the Suburban Washington Region, according to page 25 of the pdf. This region is made up of Frederick County, Prince George's County and Montgomery County. However, PG and Montgomery did not have a net positive migration over the years and it seems that both those counties negative migration overshadowed Frederick County's positive migration, since the Suburban Washington Region also has a net negative migration. Possible thought, people wanted to leave these counties and move to a more rural area of Maryland but stay relatively close to DC.  