---
title: "lab_01"
author: "James Hartner"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**). 
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**). 
* write out the answer in the form of a complete sentence in the space given (look for **A**). 
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS. 

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse.To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}


# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data 

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data. 

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data. 
**Q1** How many rows and columns does this dataframe have?
**A1** There were 33703 rows and 63 columns in this dataframe.

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

**Q2** Which West Virginia county got the most loans? How many?
**A2** Kanawha has the most loans with 4332 loans.

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** Wirt had the fewest loans with 36 loans. 

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(count_loans)
```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** [write your answer here]
According to the Census report of 2020, Kanawha County is the most populated county (183,279) in West Virginia while Wirt is the least populated (5,798). https://www.westvirginia-demographics.com/counties_by_population


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** Greenbriar Hotel Corporation received the state's largest PPP loan of $8912202. The company (corporation) is a luxury resort located in West Virginia. According to a Google search though, there is a company that is spelled as "Greenbrier" instead of what the data shows. Google did not recognized "Greenbriar" so I am assuming that Greenbriar is misspelled in the data. Anyways, the loan was able to retain 500 jobs. 

```{r}
# Write the code to produce your answer here
wv_ppp_loans   %>%
  group_by(name, amount, business_type, jobs_retained) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(amount))
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 2? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** [write your answer here] According to this article posted by ProPulica, Jim Justice owns the resort as mentioned in the previous question. 
Source-https://www.propublica.org/article/west-virginia-greenbrier-governor-jim-justice-little-trump
Another source-https://www.nationofchange.org/2020/07/07/companies-owned-by-this-billionaire-governor-received-up-to-24-million-in-bailout-loans/


**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings. 
I would drop in the average, smallest and largest loan amounts to show how much the max loan was of an outlier to the dataset. I think it would be a great investigative reporting because of who owns this resort (and his family) and why they needed such a large loan in the first place. Another nugget I would use is the total amount of all loans and see the percentage of what the max loan was in the total amount. 
**A7** [write your answer here]
```{r}
# Write the code to produce your answer here
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(max_loan_amount)
```

**Q8** Produce a summary table that has the following information for each West Virginia county:
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_county_summary` so that you can examine it with the data viewer. 

What newsworthy nuggets or patterns jump out at you that you could imagine dropping into a news story or doing follow up reporting on? Write up a few sentence that summarizes one county. 

**A8** [write your answer here]
What is really interesting is how much that loan of the resort had caused Greenbrier county to have the 12th most total amount of loans taken out. You can tell how large that loan was when Greenbrier is standing at the 13th most loans taken but has significantly less than the 12th spot (Greenbrier had 908 loans taken out whereas the county above, Marion, took out 100 more loans). Also, you can mention how much of an outlier that loan that the resort took out because of mean amount for that county (mean was 65651.80 dollars whereas the resort loan was 8912202 dollars).You could hound on this loan being astronomically insane in your story and that. Lastly, you could compare the loans of each other counties. It would be interesting to show a story about how many loans that the lower counties had could make up for the amount of that singular loan in Greenbrier. 
```{r}
# Write the code to produce your answer here
wv_county_summary <- wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(total_loans_amount))
```
