---
title: "pre_lab_03.Rmd"
author: "James Hartner"
date: "9/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 10 and 11 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab.  

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

## Instructor in class demonstration outline

Work through the code below, chapter by chapter.

## Chapter 10

### Task 1: Load libraries and settings
**Task** Run the following code in the gray-colored codeblock below to load the tidyverse library and turn off scientific notation.  

```{r}
# Remove scientific notation
options(scipen=999)
# Load the tidyverse   
library(tidyverse)
```

### Task 2: Load data
**Task** Load the PPP loan dataset by running the following code. We'll use the guess_max() function as an argument to use the first 10 rows to set the data type. What does the first line of the red Warning message that prints out when you load the data say? Answer below.
**Answer**
The first line of the red warning message prints out "One of more parsing issues see 'problems()' for details" follwed by how many rows and columns are in the data set.

```{r}
ppp_maryland_loans <- read_csv("ppp_loans_md.csv.zip", guess_max=10)

```

### Task 3: Check for problems on load
**Task** Check for problems that occurred when loading the ppp loan data for Maryland by running the following code.  How many problems were there, as shown by the number of rows in the output table showing errors? Answer below.
**Answer**
There was 87927 problems in the output table.
```{r}

problems(ppp_maryland_loans)

```

### Task 4: Reload data
**Task** Run the following codeblock to reload the data, using every row to set the data types. Does it show any parsing errors when you run? Answer below
**Answer**
It does not show any parsing errors when it runs.
```{r}
ppp_maryland_loans <- read_csv("ppp_loans_md.csv.zip")
```

### Task 5: Examine the data with glimpse
**Task** Run the following codeblock to glimpse the data.  What data type is the "amount" field?  What data type is the "name" field? What data type is the id column? Answer below.
**Answer**
The "name" field has a <chr> id.
```{r}
glimpse(ppp_maryland_loans)
```
### Task 6: Change the "id" field data type
**Task** Run the following codeblock to change the data type of the "id" field from a number (dbl) to a character, and then glimpse the data, to see the change.     

```{r}
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))

glimpse(ppp_maryland_loans)
```

### Task 7: Examine the franchise column for missing values.
**Task** Run the following codeblock to group by franchise_name, count the numer of loans, then sort from highest to lowest on count. How many loans are there where franchise_name is NA? How many loans are there where the franchise_name is "Subway"? Answer below.
**Answer**:
There were 192959 loans in the NA section and 275 loans for Subway.

```{r}

ppp_maryland_loans %>%
  group_by(franchise_name) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```
### Task 8: Examine the forgiveness_amount column for missing values.
**Task** Run the following codeblock to group by forgiveness_amount, count the numer of loans, then sort from highest to lowest on count. How many loans are there where forgiveness_amount is NA? Answer below.
**Answer**:
There were 135073 loans in the forgiveness_amount.
```{r}

ppp_maryland_loans %>%
  group_by(forgiveness_amount) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```

### Task 9: Examine the forgiveness_date column for missing values.
**Task** Run the following codeblock to group by forgiveness_date, count the numer of loans, then sort from highest to lowest on count and filter to only show NA values. How many loans are there where forgiveness_date is NA? Is it the same as the number for forgiveness_amount? Answer below.
**Answer**:
There were 135073 loans in the forgiveness_date, which is the same as forgiveness_amount.

```{r}

ppp_maryland_loans %>%
  group_by(forgiveness_date) %>%
  summarise(
    count=n()
  ) %>%
  filter(is.na(forgiveness_date)) %>%
  arrange(desc(count))
```
### Task 10: Examine forgiveness_date and forgiveness_amount together
**Task** Run the following codeblock to group by forgiveness_date and forgiveness_amount, count the number of loans, then sort from highest to lowest on count and filter to only show NA values. How many loans are there where forgiveness_date and forgiveness_amount is NA? Is it the same as the number for forgiveness_amount and forgiveness_amount run separately in the codeblocks above? What can you conclude from Answer below.
**Answer**:
The exact same number of 135073 appear. I think that R is taking all of the business that didn't list a franchise in the program and grouping it together as both forgiveness_amount and forgiveness_date. We also aren't looking at the specific amounts or dates, we are just pulling how many businesses didn't list themselves as a franchise in the loans. 
```{r}
ppp_maryland_loans %>%
  group_by(forgiveness_amount, forgiveness_date) %>%
  summarise(
    count=n()
  ) %>%
  filter(is.na(forgiveness_date)) %>%
  arrange(desc(count))

```
### Task 11: Install lubridate
**Task** Run the following codeblock to install the lubridate package.
```{r}
install.packages('lubridate')
```

### Task 12: Load lubridate
**Task** Run the following code to load the lubridate library.

```{r}
library(lubridate)
```

### Task 13: Look for date gaps in data
**Task** Run the following codeblock to create a new column called "month_year_approved" that extracts the month and year from the "date_approved" column.  Group by the new "month_year_approved" column, count, and sort by "month_year_approved". How many loans are there in the data for October 2020? Answer below.

**Answer**
It claims that there were 0 loans in the data for October 2020.
```{r}
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)
```

### Task 14: Check for suspicious outliers
**Task** Run the following codeblock to find the largest and smallest loans in the data, by using summarise() without grouping. How much is the largest loan for? How much is the smallest loan for? Are either of these suspicious to you? Why or why not? Answer below.
The maximum loan is 10000000 dollars and the smallest loan is $6. I think the six dollar one is more suspicous because who would take out a loan of only six dollars? I feel like it is a possible data entry to skew the data a bit. 
**Answer**
```{r}

ppp_maryland_loans %>%
  summarise(max_amount = max(amount),
            min_amount= min(amount))
```


### Task 15: Look at smallest loans
**Task** Run the following codeblock to examine all loans under $100. How many are there? Answer below.
**Answer**
There were two loans under $100.
```{r}

ppp_maryland_loans %>%
 filter(amount < 100)
```

### Task 16: Look for internal consistency
**Task** Run the following codeblock to examine all loans under $100 and look at whether the various fields with amounts align. Do they? Answer below.
**Answer**
They all align to exactly the same numbers in each field. I feel like the payroll_proceed shouldn't be the same numbers though because of the fact the payroll should make up 60% of it. I'm surprised that the other 40% was used for payroll, but at the same time, it's a little suspicous with the amount of money that they took out. 
```{r}

ppp_maryland_loans %>%
 filter(amount < 100) %>%
  select(name, amount, initial_approval_amount, current_approval_amount, payroll_proceed)
```

## Chapter 11

### Task 1: Install janitor
**Task** Run the following codeblock to install the janitor package.
```{r}
install.packages('janitor')
```

### Task 2: Load janitor and the tidyverse
**Task** Run the following code to load the tidyverse and janitor.

```{r}
library(tidyverse)
library(janitor)
```

### Task 3: Load Arnold, Maryland demonstration data
**Task** Run the following codeblock to load a demonstration slice of the data for Arnold, Maryland. How many rows are in this demonstration data set?
**Answer**
There are six rows in the demonstatrion data set. 
```{r}
arnold_md_loans <- read_rds("arnold_md_loans.rmd")
```

### Task 4: Examine the data with glimpse
**Task** Run the following codeblock to glimpse the data.  What data type is the "amount" field? Answer below.
**Answer**
The "amount" field has a data type of <chr>.
```{r}
glimpse(arnold_md_loans)
```

And let's examine the full data set.

### Task 5: Examine the data table
**Task** Run the following codeblock to examine the data. Name three problems the book chapter says exist in this data set that could prevent us from answering questions?  Answer below
**Answer**
With this data, the amount is stored as a character, the zip code had five digits in the bottom three rows and then on the top and for some of the rows, Arnold is misspelt or all caps.
```{r}
arnold_md_loans
```

### Task 6: Use clean_names()
**Task** Run the following codeblock to use the clean_names() function from janitor to standardize column names. How does the function change the name of the column "1_id"? Answer below.
**Answer**
The name changes to "x1_id."

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names()

# display the cleaned dataset
cleaned_arnold_md_loans
```

### Task 7: Use rename()
**Task** Run the following codeblock to use the clean_names() function from janitor to standardize column names and then use rename() to change the "x1_id" column.


```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id)

# display the cleaned dataset
cleaned_arnold_md_loans


```

### Task 8: Try summarizing the amount column
**Task** Run the following codeblock to attempt to add up the amount of all loans. What does the error say when you run this code? What do you think it means? Answer below.
**Answer**
It says, "Error: Problem with `summarise()` column `total_amount`.
i `total_amount = sum(amount)`.
x invalid 'type' (character) of argument" and tells us to run 'rlang::last_error() to see where the error occurred. I think the characters are not the same or the character in "amount" is not an integer, which means we can't add those numbers up. 
```{r}
# cleaning function
total_arnold_md_loans <- cleaned_arnold_md_loans %>%
  summarise(total_amount = sum(amount))

# display the cleaned dataset
total_arnold_md_loans


```
### Task 9: Change data type for amount
**Task** Run the following codeblock to attempt to change the datatype for the amount field to a number. What is the new data type (three letter code) for amount? Answer below.
**Answer**
The new data type for amount is <dbl>.
```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount))


# display the cleaned dataset
cleaned_arnold_md_loans

```
### Task 10: Try again to summarize the amount column
**Task** Run the following codeblock to add up the amount of all loans in this data.  What is the total? Answer below.
**Answer**
The amount of all loans is $2,033,179.
```{r}
# cleaning function
total_arnold_md_loans <- cleaned_arnold_md_loans %>%
  summarise(total_amount = sum(amount))

# display the cleaned dataset
total_arnold_md_loans


```

### Task 11: Check for duplicate rows
**Task** Run the following codeblock to check for duplicate rows using get_dupes().  How many are there? What is the business name? Answer below.
**Answer**
There are 2 duplicate rows. The business name is "Stuart Jones LLC."

```{r}
cleaned_arnold_md_loans %>%
  get_dupes()
```

### Task 12: Check for duplicate rows
**Task** Run the following codeblock to use distinct() to get rid of duplicate rows. How many rows does the new dataframe have? Answer below.
**Answer**
The new dataframe only has five rows now.
```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct()


# display the cleaned dataset
cleaned_arnold_md_loans

```

### Task 13: Clean up ZIP code
**Task** Run the following codeblock to use str_sub() to convert the ZIP codes that have nine digits to five digits, standardizing the field.
**Answer**
The zip is 21012.
```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))


# display the cleaned dataset
cleaned_arnold_md_loans

```

### Task 14: Clean up city field
**Task** Run the following codeblock to use str_tot_title() to standarize capitalization in the "city" field. How many mispellings of Arnold remain after running this code? Answer below.
There are two misspellings of Arnold remaining.
```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city))


# display the cleaned dataset
cleaned_arnold_md_loans

```


### Task 15: Clean up city field more with case_when()
**Task** Run the following codeblock to use case_when() to fix misspellings of Arnold in the "city" field. How many mispellings of Arnold remain after running this code? Answer below.
**Answer**
There is now only one misspelling. 

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    TRUE ~ city
  ))


# display the cleaned dataset
cleaned_arnold_md_loans

```

### Task 16: Clean up city field more with case_when()
**Task** Run the following codeblock to use case_when() to fix misspellings of Arnold in the "city" field using both the exact match method and the str_detect() method. How many mispellings of Arnold remain after running this code? Answer below.
**Answer**
There are no misspellings of Arnold now.
```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    str_detect(city,"^Arnl") ~ "Arnold",
    TRUE ~ city
  ))


# display the cleaned dataset
cleaned_arnold_md_loans

```
